---
title: "p8105_hw2_st3117"
author: "Sha Tao"
date: "October 6, 2018"
output: github_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges)
library(patchwork)
library(hexbin)
library(p8105.datasets)

```

## Problem 1_1 BRFSS Data Cleaning

```{r Problem 1_1}

brfss = 
  brfss_smart2010 %>% 
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>% 
  separate(locationdesc, into = c("remove2", "county"), sep = 5) %>% 
  select(year, state = locationabbr, county, response, data_value) %>% 
  mutate(response = as.factor(response),
         response = factor(response, levels = c("Excellent", "Very good", "Good", "Fair", "Poor")))

levels(brfss$response)

brfss

```

## Problem 1_2 Answering Questions

```{r Problem 1_2}

brfss %>% 
  spread(key = response, value = data_value) %>% 
  janitor::clean_names() %>% 
  filter(year == 2002) %>% 
  group_by(state) %>% 
  summarize(n = n()) %>% 
  filter(n == 7)

brfss %>% 
  filter(year == 2002) %>% 
  group_by(state) %>% 
  summarize(n = n_distinct(county)) %>% 
  filter(n == 7)


```

## Problem 1_3 Spaghetti Plot showing the number of observations in each state from 2002 to 2010s

```{r Problem 1_3}

brfss %>%
  group_by(year, state) %>% 
  summarize(n_obs = n()) %>% 
  ggplot(aes(x = year, y = n_obs, color = state)) +
    geom_line() +
    labs(
      title = "Number of Observations Plot",
      x = "Year",
      y = "Number of Observations"
      ) +
    theme_bw() +
    theme(legend.position = "right")

```

## Problem 1_4 NY State “Excellent” responses

```{r Problem 1_4}

brfss %>%
  filter(state == "NY" & response == "Excellent") %>% 
  filter(year == 2002 | year == 2006 | year == 2010) %>%
  group_by(county) %>%
  summarize(mean_prop = mean(data_value, na.rm = TRUE),
            sd_prop = sd(data_value, na.rm = TRUE)) %>% 
  knitr::kable(digits = 3)

```

## Problem 1_5  Five-panel plot showing the distribution of state-level averages over time

```{r Problem 1_5}

brfss %>% 
  group_by(year, state, response) %>% 
  summarize(mean_prop = mean(data_value, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = mean_prop, color = state)) +
    geom_line() +
    facet_grid(. ~ response) +
    labs(
      title = "Distribution of State-level Response over time",
      x = "Year",
      y = "Mean Response Proportion"
      ) +
    theme(axis.text.x = element_text(hjust = 1, angle = 45))

```

## Problem 2_2  Five-panel plot showing the distribution of state-level averages over time

```{r Problem 2_2}

instacart = instacart

## How many aisles are there, and which aisles are the most items ordered from?

count(distinct(instacart, aisle_id)) 
  
instacart %>%
  summarize(n = n_distinct(aisle_id))

instacart %>%
  group_by(aisle) %>%
  summarize(n_obs = n()) %>% 
  arrange(-n_obs) %>% 
  filter(row_number() == 1)

```

## Problem 2_3 Make a plot that shows the number of items ordered in each aisle

```{r Problem 2_3, fig.width = 10, fig.height = 20}

instacart %>%
  group_by(aisle) %>%
  summarize(n_obs = n()) %>% 
  mutate(group = as.numeric(cut_number(n_obs, 3))) %>% 
  ggplot(aes(x = reorder(aisle, n_obs), y = n_obs)) +
    geom_point() +
    facet_wrap(group ~ ., nrow = 3, scales = "free") +
    theme(axis.text.x = element_text(size = 8, hjust = 1, angle = 45)) +
    labs(
      title = "Number of Items Ordered in Aisles",
      x = "Aisle Name",
      y = "Number of Items Ordered"
      )

```

## Problem 2_4 Make a table showing the most popular item in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”

```{r Problem 2_4}

instacart %>% 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>% 
  group_by(aisle, product_name) %>% 
  summarize(n_obs = n()) %>%
  summarize(max_product = max(n_obs))
            
instacart %>% 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>% 
  group_by(aisle, product_name) %>% 
  summarize(n_obs = n()) %>% 
  arrange(aisle, -n_obs) %>% 
  filter(row_number() == 1) %>% 
  knitr::kable()

```

## Problem 2_5 Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)

```{r Problem 2_5}

instacart %>% 
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day, na.rm = TRUE)) %>% 
  spread(key = product_name, value = mean_hour) %>% 
  knitr::kable(digits = 2)

```

## Problem 3_1

```{r Problem 3_1}

ny_noaa = ny_noaa

ny_noaa = ny_noaa %>% 
  mutate(year = strftime(date, "%Y"),
         month = strftime(date, "%B"),
         day = strftime(date, "%d"),
         prcp = prcp / 10,
         snow = as.numeric(snow),
         snwd = as.numeric(snwd),
         tmax = as.numeric(tmax) / 10,
         tmin = as.numeric(tmin) / 10) %>% 
  select(id, year, month, day, prcp_mm = prcp, snow_mm = snow, snwd_mm = snwd, tmax_c = tmax, tmin_c = tmin)



```

## Problem 3_2

```{r Problem 3_2}

ny_noaa %>% 
  group_by(snow_mm) %>% 
  summarize(n_obs = n()) %>% 
  arrange(-n_obs) %>% 
  head(1)

ny_noaa %>% 
  group_by(month) %>% 
  summarize(mean_snow = median(snow_mm, na.rm = TRUE))


```

## Problem 3_3 Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?

```{r Problem 3_3, out.width="800px", out.height="800px"}

ny_noaa %>% 
  filter(month == "January" | month == "July") %>% 
  group_by(id, month) %>% 
  summarize(mean_tmax = mean(tmax_c, na.rm = TRUE)) %>% 
  filter(!is.na(mean_tmax)) %>% 
  ggplot(aes(x = id, y = mean_tmax)) +
    geom_point() +
    facet_grid(month ~ .) +
    theme(axis.text.x = element_text(size = 4, hjust = 1, angle = 45)) +
    labs(
      title = "Average Max Temperature in January and in July",
      x = "Station ID",
      y = "Average Max Temperature"
      )

ny_noaa %>% 
  filter(month == "January" | month == "July") %>% 
  group_by(id, month) %>% 
  summarize(mean_tmax = mean(tmax_c, na.rm = TRUE)) %>% 
  filter(!is.na(mean_tmax)) %>% 
  ggplot(aes(x = month, y = mean_tmax)) +
    geom_boxplot()

is_outlier = function(x) {
  return(x < quantile(x, 0.25) - 1.5 * IQR(x) | x > quantile(x, 0.75) + 1.5 * IQR(x))
}

ny_noaa %>% 
  filter(month == "January" | month == "July") %>% 
  group_by(id, month) %>% 
  summarize(mean_tmax = mean(tmax_c, na.rm = TRUE)) %>% 
  filter(!is.na(mean_tmax)) %>% 
  group_by(month) %>% 
  mutate(outlier = ifelse(is_outlier(mean_tmax), mean_tmax, as.numeric(NA))) %>% 
  filter(!is.na(outlier)) %>% 
  arrange(month)
    

```

## Problem 3_4 Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.


```{r Problem 3_4}

tmax_tmin_p = ggplot(ny_noaa, aes(x = tmax_c, y = tmin_c)) + 
                geom_hex() +
                labs(
                  title = "tmax vs tmin",
                  x = "tmax",
                  y = "tmin"
                  )

```

## Problem 3_5 and 
```{r Problem 3_5}

snow_dist_p = ny_noaa %>% 
                filter(snow_mm > 0 & snow_mm < 100) %>% 
                ggplot(aes(x = year, y = snow_mm)) +
                  geom_violin() +
                  theme(axis.text.x = element_text(size = 8, hjust = 1, angle = 45)) +
                  labs(
                    title = "Distribution of Snowfall",
                    x = "Year",
                    y = "Snowfall in mm"
                    )

```

```{r}

tmax_tmin_p / snow_dist_p

```


