---
title: "p8105_hw2_st3117"
author: "Sha Tao"
date: "October 6, 2018"
output: github_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges)
library(patchwork)
library(p8105.datasets)

```

## Problem 1_1 BRFSS Data Cleaning

```{r Problem 1_1}

brfss = 
  brfss_smart2010 %>% 
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>% 
  separate(locationdesc, into = c("remove2", "county"), sep = 5) %>% 
  select(year, state = locationabbr, county, response, data_value) %>% 
  mutate(response = as.factor(response),
         response = factor(response, levels = c("Excellent", "Very good", "Good", "Fair", "Poor")))

levels(brfss$response)

brfss

```

## Problem 1_2 Answering Questions

```{r Problem 1_2}

brfss %>% 
  spread(key = response, value = data_value) %>% 
  janitor::clean_names() %>% 
  filter(year == 2002) %>% 
  group_by(state) %>% 
  summarize(n = n()) %>% 
  filter(n == 7)

brfss %>% 
  filter(year == 2002) %>% 
  group_by(state) %>% 
  summarize(n = n_distinct(county)) %>% 
  filter(n == 7)


```

## Problem 1_3 Spaghetti Plot showing the number of observations in each state from 2002 to 2010s

```{r Problem 1_3}

brfss %>%
  group_by(year, state) %>% 
  summarize(n_obs = n()) %>% 
  ggplot(aes(x = year, y = n_obs, color = state)) +
    geom_line() +
    labs(
      title = "Number of Observations Plot",
      x = "Year",
      y = "Number of Observations"
      ) +
    theme_bw() +
    theme(legend.position = "right")

```

## Problem 1_4 NY State “Excellent” responses

```{r Problem 1_4}

brfss %>%
  filter(state == "NY" & response == "Excellent") %>% 
  filter(year == 2002 | year == 2006 | year == 2010) %>%
  group_by(county) %>%
  summarize(mean_prop = mean(data_value, na.rm = TRUE),
            sd_prop = sd(data_value, na.rm = TRUE)) %>% 
  knitr::kable(digits = 3)

```

## Problem 1_5  Five-panel plot showing the distribution of state-level averages over time

```{r Problem 1_5}

brfss %>% 
  group_by(year, state, response) %>% 
  summarize(mean_prop = mean(data_value, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = mean_prop, color = state)) +
    geom_line() +
    facet_grid(. ~ response) +
    labs(
      title = "Distribution of State-level Response over time",
      x = "Year",
      y = "Mean Response Proportion"
      ) +
    theme(axis.text.x = element_text(hjust = 1, angle = 45))

```

## Problem 2_1  Five-panel plot showing the distribution of state-level averages over time

```{r Problem 2_1}

instacart = instacart

## How many aisles are there, and which aisles are the most items ordered from?

count(distinct(instacart, aisle_id)) 
  
instacart %>%
  summarize(n = n_distinct(aisle_id))

instacart %>%
  group_by(aisle) %>%
  summarize(n_obs = n()) %>% 
  arrange(-n_obs) %>% 
  filter(row_number() == 1)

```

## Problem 2_2 Make a plot that shows the number of items ordered in each aisle

```{r Problem 2_2, fig.width = 10, fig.height = 20}

instacart %>%
  group_by(aisle) %>%
  summarize(n_obs = n()) %>% 
  mutate(group = as.numeric(cut_number(n_obs, 3))) %>% 
  ggplot(aes(x = reorder(aisle, n_obs), y = n_obs)) +
    geom_point() +
    facet_wrap(group ~ ., nrow = 3, scales = "free") +
    theme(axis.text.x = element_text(size = 8, hjust = 1, angle = 45)) +
    labs(
      title = "Number of Items Ordered in Aisles",
      x = "Aisle Name",
      y = "Number of Items Ordered"
      )

```

## Problem 2_3 Make a table showing the most popular item in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”

```{r Problem 2_3}

instacart %>% 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>% 
  group_by(aisle, product_name) %>% 
  summarize(n_obs = n()) %>%
  summarize(max_product = max(n_obs))
            
instacart %>% 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>% 
  group_by(aisle, product_name) %>% 
  summarize(n_obs = n()) %>% 
  arrange(aisle, -n_obs) %>% 
  filter(row_number() == 1) %>% 
  knitr::kable()

```

## Problem 2_4 Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)

```{r Problem 2_4}

instacart %>% 
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day, na.rm = TRUE)) %>% 
  spread(key = product_name, value = mean_hour) %>% 
  knitr::kable(digits = 2)

```






